---
title: "Day 7 Practice"
output: html_document
---

In this practice, let's explore data on tipping, in the `tips` data frame from the `reshape2` package.

0. First get a little familiar with the structure of the data, using `summary()`, `head()`, etc.  What are the numeric variables?  What are their units? What about factors? Are factor levels sensible?

```{r}

```

1. Create a simple scatterplot, plotting the tip amount by the total bill.
  - first plot the raw quantities
  - "save" the initial simple plot as an object
  - change the axis labels for both x and y
    - try two different methods
  
```{r}

```

2. Continue using the previous data
  - using `scales`, plot the axes so that they follow a logarithmic pattern
  - add "break" values every 5 dollars on the x axis, and every 2 dollars on the y axis
  - finally, use `labels = dollar` (made available by the `scales` package) to format the axis numbers as dollars

```{r}

```

3. Continue to build off of the final scatterplot in the previous problem (hint: you may need to change the plot object, in order to make the following changes)
  - map gender to the color of the points
  - change the color values, using both the "brewer" and "manual" methods
  - add regression lines (regular linear regression), which also break down by gender
  - also try a loess smoother instead of regression lines
  - finally, change the overall "theme" to either the "classic", "minimal", or "bw" theme

```{r}

```


4. Now let's modify the legend for the `sex` variable, building off off the scatterplot with linear regression lines
  - using the appropriate `scale` function, change the name of the legend to "Gender"
  - change the location of the legend with the `theme()` function
  - also using the "legend.position" argument, remove the legend completely
  - finally, put the legend back to the default position, but change the levels, such that Male is given as the first level
  
```{r}

```

5. Let's explore how this pattern might vary for different days of the week.
  - use faceting to create a separate plot for each day
  - try both the "wrap" and "grid" faceting functions
  - make a change so that the labels on the top of the facets give the full day name (e.g., "Friday", not "Fri"), and they are ordered from Thursday to Sunday
  
```{r}


```

6. Now let's look at adding information about the size of the party to the plot
  - first leave off the regression lines, to make it easier to see
  - map party size to the size of the points
  - look at both an aggregate plot and plots faceted by day
  - now add back the regression lines, but only plot a single regression line (i.e., not broken down by gender), remove the confidence interval (hint: look at the arguments for `geom_smooth()`), and make the line a dashed black line
  
```{r}

```

