---
title: "Day 7 Homework"
output: html_document
---

Find three relationships in your data.  For each relationship, do the following:

1. Verbally articulate the relationship you want to examine with a plot.

2. Verbally identify what variables you want to map to each graphical dimension.

a) relationship between WithinMinPairs and Merged for all languages (Corpus)
b) correlations between AllEntropy and each other Entropy variable
c) relationship between the number of sound pairs in a language and the the average number of minimal pairs


3. Reshape your data as needed.

```{r}
scottdata <- read.csv("scottdata/CognitionPaperFinalData.csv")
scottdata.long <- gather(scottdata, variable, value, WithinEntropy:BiphoneEntropy)
pairs.sum <- group_by(scottdata, Corpus) %>% summarize(total.pairs = length(Pair), mean.minpair = mean(AllMinPairs))

```

4. Use ggplot2 to create your graphs.

```{r}
ggplot(scottdata, aes(Corpus, log(WithinMinPairs+1))) + geom_boxplot(aes(fill = Merged)) + scale_fill_brewer(palette = "Set1")

ggplot(scottdata.long, aes(value, AllEntropy)) + geom_point() + geom_smooth(method = "lm") + facet_wrap(~ variable, scales = "free")

ggplot(pairs.sum, aes(total.pairs, mean.minpair)) + geom_text(aes(label = Corpus))

```

5. Fiddle with colors, axis labels, themes, etc. until the plot looks "publication ready" to you.  Try to change at least one thing per plot.

6. Save all plots to a format that you could use in a paper.  HINT: PNG works well for pasting into Word documents.

```{r}
png("homework_plot1.png", width = 10, height = 4, units = "in", res = 300)
ggplot(scottdata, aes(Corpus, WithinMinPairs)) + geom_boxplot(aes(fill = Merged)) + scale_fill_brewer(palette = "Set1", name = "") + theme_minimal() + scale_y_log10(name = "within-category counts of minimal pairs", breaks = c(5, 10, 25, 50, 100, 250, 500, 1000))
dev.off()

png("homework_plot2.png", width = 10, height = 4, units = "in", res = 300)
ggplot(scottdata.long, aes(value, AllEntropy)) + geom_point(alpha = .3) + geom_smooth(method = "lm") + facet_wrap(~ variable, scales = "free") + theme_minimal() 
dev.off()

png("homework_plot3.png", width = 5, height = 5, units = "in", res = 300)
ggplot(pairs.sum, aes(total.pairs, mean.minpair)) + geom_text(aes(label = Corpus)) + theme_minimal() + xlab("total count of sound pairs examined") + ylab("average count of minimal pairs")
dev.off()
```


```{r}
png("")
```

