---
title: "Day 8 Lecture"
output: html_document
---

# Review HW 7

# Loops and control
  - if/else
  - for-loops
    - counters (sets)
    - "growing" results
    - debugging
    - break, next
  - simulating data
  - "roll your own" power analysis

# if/else
  - form: if(LOGICAL TEST) { CODE BLOCK if TEST == TRUE  }
  -       else if(OPTIONAL ADDITIONAL TEST) { CODE BLOCK if TRUE }   <-- "else" code only tested if initial "if" fails

```{r}
x <- 2
if(x > 1){
  print(x)
}

x <- 2
if(x > 1) {
  print("you have enough!")
} else {
  print("you need more!")
}

x <- 3
if(x > 3) {
  print("you have too many!")
} else if(x < 3) {
  print("you need more!")
} else {
  print("just right!")
}

x <- c(1:10)
if(anyNA(x)) {
  number.of.NAs <- sum(is.na(x))
  print(paste("The mean (ignoring", number.of.NAs, 
              "missing values) is", mean(x, na.rm = TRUE)))
} else {
  print(paste("The mean is", mean(x)))
}

```

  - `ifelse()`
    - arguments are: test = [LOGICAL TEST], yes = [VALUE if TEST == TRUE], no = [VALUE if TEST == FALSE]

```{r} 
x <- 4
result <- ifelse(x > 3, "yay", "oh no!")
result

x <- 3
result <- ifelse(x > 3, "yay", ifelse(x < 3, "oh no!", "well, okay"))
result

# redux
x <- c(1:10, NA, NA)
if(anyNA(x)) {
  number.of.NAs <- sum(is.na(x))
  print(paste("The mean (ignoring", number.of.NAs, 
              "missing", ifelse(number.of.NAs == 1, "value)", "values)"), "is", mean(x, na.rm = TRUE)))
} else {
  print(paste("The mean is", mean(x)))
}

```

# PRACTICE #1 & #2

# `for` loops
  - simple way to iterate through a set of code, via a "counter" variable
  - avoid re-writing code, when possible!

```{r}
mydf <- data.frame(x = rnorm(n = 1000000, mean = 2, sd = 1), group = rep(1:1000, each = 1000))

rep(1:3, 3)
rep(1:3, each = 3)
rep(1:3, each = 3, 2)

this.vector <- mydf[mydf$group == 1, "x"]
my.answer1 <- mean(this.vector)
this.vector <- mydf[mydf$group == 2, "x"]
my.answer2 <- mean(this.vector)
this.vector <- mydf[mydf$group == 3, "x"]
my.answer3 <- mean(this.vector)

# MAKE IT A LOOP

for(my.counter in 1:10) {
  print(my.counter * 2)  
}

for(i in c(9, 4, 25)) {
  print(i * 2)  
}

for(this.name in c("Ally", "Bob", "Carol")) {
  print(this.name)
}

# WARNING: you must create objects before you can subset them
# my.answer <- vector()
for(this.group in c(1, 2, 3)) {
  this.vector <-  mydf$x[mydf$group == this.group]
  my.answer[this.group] <- mean(this.vector)  
}
my.answer

# Will work, but may get slow (in theory)
my.answer <- NA
system.time(
for(this.group in 1:1000) {
  this.vector <-  mydf[mydf$group == this.group, "x"]
  my.answer[this.group] <- mean(this.vector)
})
length(my.answer)

# BEST PRACTICE: pre-create the "results" object, with the proper size,
#                to be "filled in" by the loop
my.groups <- 1:1000

my.answer <- vector(length = length(my.groups))

system.time(for(this.counter in 1:length(my.groups)) {
  this.group <- my.groups[this.counter]
  this.vector <- mydf[mydf$group == this.group, "x"]
  my.answer[this.counter] <- mean(this.vector)
})
length(my.answer)
mean(my.answer)

```

# simulating data
  - d/p/q/r* family of functions for different distributions

```{r}
?rnorm

x <- sort(rnorm(1000))
summary(x)
sd(x)
hist(x)
plot(x)
plot(dnorm(x) ~ x)
plot(pnorm(x) ~ x)
plot(qnorm(seq(0, 1, .01)) ~ seq(0, 1, .01))

y <- rnorm(1000, mean = 1, sd = 2)
summary(y)
sd(y)

library(MASS) # look out for "masking" of select() from dplyr

cor.table <- matrix(c(1, .2, .2, 1), nrow = 2)
cor.table
some.data <- mvrnorm(n = 10, mu = c(1, 2), Sigma = cor.table, empirical = FALSE)
some.data
cor(some.data[, 1], some.data[, 2])
as.data.frame(some.data)
```

# Combining sims and loops

```{r}
mu.sigmas <- data.frame(mean1 = c(0, 0.1, -0.1, -0.3, 0.25), mean2 = c(0.2, 0.0, 0.3, -0.15, -0.1), corr = c(.2, .4, .1, .35, -0.3))
mu.sigmas

data.sets <- list()
length(data.sets) <- 5
for(i in 1:length(data.sets)) {
  data.sets[[i]] <- as.data.frame(mvrnorm(n = 1000, mu = c(mu.sigmas[i, "mean1"], mu.sigmas[i, "mean2"]), Sigma = matrix(c(1, mu.sigmas[i, "corr"], mu.sigmas[i, "corr"], 1), ncol = 2)))
  data.sets[[i]]$set <- i
  write.csv(data.sets[[i]], file = paste("sampledata", i, ".csv", sep = ""), row.names = FALSE)
}

str(data.sets)
library(dplyr)
all.data.sets <- bind_rows(data.sets, .id = "setID")
xtabs(~ set + setID, all.data.sets)

a1 <- 1
a2 <- 55
a3 <- 43
a4 <- 1001

# get() function
for(this.var in paste("a", 1:4, sep = "")) {
  print(get(this.var) * 2)
}


```

# Power analysis

```{r}
# set up the parameters of the simulation
n.sim <- 10000
sample.n <- 100
correlation <- .2

cortable <- matrix(c(1, correlation, correlation, 1), ncol = 2)
# "pre-allocate" the results table
results <- data.frame(iteration = 1:n.sim, p = NA)

# run a repeated simulation of your planned analysis
for(counter in 1:n.sim) {
  this.data <- as.data.frame(mvrnorm(sample.n, mu = c(0, 0), Sigma = cortable))
  this.cortest <- cor.test(this.data$V1, this.data$V2)
  results[counter, "p"] <- this.cortest$p.value
  # "poor man's progress bar"
  if(counter %% 500 == 0) {
    print(paste("finished simulation #", counter))
  }
}
head(results)
results$sig <- ifelse(results$p < .05, 1, 0)
head(results)
mean(results$sig)
```

Going further: looping over sets of simulations!

```{r}
# set up the parameters of the simulation
n.sim <- 10000
sample.ns <- seq(100, 500, 50)
correlation <- .2

power.analysis.results <- data.frame(sample.size = sample.ns, power = NA)

for(this.sample in sample.ns) {
  cortable <- matrix(c(1, correlation, correlation, 1), ncol = 2)
  # "pre-allocate" the results table
  results <- data.frame(iteration = 1:n.sim, p = NA)
  
  for(counter in 1:n.sim) {
    this.data <- as.data.frame(mvrnorm(this.sample, mu = c(0, 0), Sigma = cortable))
    this.cortest <- cor.test(this.data$V1, this.data$V2)
    results[counter, "p"] <- this.cortest$p.value
    if(counter %% 500 == 0) {
      print(paste("finished simulation #", counter, "for sample size:", this.sample))
    }
  }
  #head(results)
  results$sig <- ifelse(results$p < .05, TRUE, FALSE)
  #head(results)
  power.analysis.results[power.analysis.results$sample.size == this.sample, "power"] <-   sum(results$sig)/nrow(results)

}

head(power.analysis.results)

power.threshold <- .9

min.sample <- min(power.analysis.results$sample.size[power.analysis.results$power >= power.threshold])
ggplot(power.analysis.results, aes(sample.size, power)) + geom_line() + 
  geom_hline(yintercept = max(power.analysis.results$power), color = "red") +
  annotate("text", x = min(power.analysis.results$sample.size) + (max(power.analysis.results$sample.size) - min(power.analysis.results$sample.size))/2, y = 1.02, label = paste("max power in these simulations:", max(power.analysis.results$power))) +
  geom_hline(yintercept = power.threshold, color = "red", linetype = 2) + 
  annotate("text", x = min(power.analysis.results$sample.size) + (max(power.analysis.results$sample.size) - min(power.analysis.results$sample.size))/2, y = power.threshold + .02, label = paste("sample needed for", power.threshold, " power:", min.sample))
    

```

