---
title: "Day 8 Practice"
output: html_document
---

1. Write an if...else statement to test whether a string matches "right", and if so, print "you win!", but if the string matches "wrong", print "you lose!", and if it's neither, print "the result is inconclusive".

```{r}

```

2. Make a copy of the iris data. Use an ifelse statement to create a new column, where:
  - if the value of Sepal.Length is less than 5, the new column should have a value of "low"
  - if the value of Sepal.Length is greater than 6, it should have a value of "high"
  - and if it's between 5 and 6 (inclusive), it should have a value of "mid"
  - make a plot in ggplot2 to confirm that you did this right
  
```{r}

```

3. Consider the iris data again.
  - make a scatterplot of Sepal.Length by Sepal.Width
  - facet the plot by Species
  - using a loop, loop through species and create three separate plots
    - in each plot, display the species as the plot title (either use `labs()` or `ggtitle()`)
    - use pdf() and dev.off() so that the three plots get written out to a single file
    
```{r}

```

4. Similar to what we did with the correlation power simulations, write code (including a loop) that creates a power simulation for comparing the means of two normally-distributed sample.
  - compare the sample with both `t.test()` and `wilcox.test()`
  - put the p-values from both tests in each simulation result
  - compare the power of the t-test vs. the power of the Wilcoxon test
  
```{r}
# set up the parameters of the simulation


# set up the "empty" data frame for the results

# start the loop
# each iteration of the loop should:
# 1. generate data
# 2. perform the analysis
# 3. store the results (here, p-values) in the results data frame


# examine the results
# use ifelse to create a column that indicates which of the p-values reached significance


# power = the proportion of "significant" results


```

