---
title: "Day 8 Practice"
output: html_document
---

1. Write an if...else statement to test whether a string matches "right", and if so, print "you win!", but if the string matches "wrong", print "you lose!", and if it's neither, print "the result is inconclusive".

```{r}
my.string <- "wrong"

if(my.string == "right") { 
  print("you win!") 
} else if(my.string == "wrong") {
  print("you lose!")
} else {
  print("the result is inconclusive")
}

```

2. Make a copy of the iris data. Use an ifelse statement to create a new column, where:
  - if the value of Sepal.Length is less than 5, the new column should have a value of "low"
  - if the value of Sepal.Length is greater than 6, it should have a value of "high"
  - and if it's between 5 and 6 (inclusive), it should have a value of "mid"
  - make a plot in ggplot2 to confirm that you did this right
  
```{r}
my.iris <- iris
my.iris$threshold <- ifelse(my.iris$Sepal.Length < 5, "low", ifelse(my.iris$Sepal.Length > 6, "high", ifelse(my.iris$Sepal.Length >= 5 & my.iris$Sepal.Length <= 6, "mid", NA)))
my.iris$threshold <- factor(my.iris$threshold, levels = c("low", "mid", "high"))

ggplot(my.iris, aes(threshold, Sepal.Length)) + geom_point()
ggplot(my.iris, aes(Sepal.Length)) + geom_histogram(aes(fill = threshold))

```

3. Consider the iris data again.
  - make a scatterplot of Sepal.Length by Sepal.Width
  - facet the plot by Species
  - using a loop, loop through species and create three separate plots
    - in each plot, display the species as the plot title (either use `labs()` or `ggtitle()`)
    - use pdf() and dev.off() so that the three plots get written out to a single file
    
```{r}
my.iris <- iris
ggplot(my.iris, aes(Sepal.Length, Sepal.Width)) + geom_point()
ggplot(my.iris, aes(Sepal.Length, Sepal.Width)) + geom_point() + facet_wrap(~ Species)

ggplot(my.iris[my.iris$Species == "virginica", ], aes(Sepal.Length, Sepal.Width)) + geom_point() + ggtitle("virginica")

pdf("loop_plots.pdf")
for(this.species in levels(my.iris$Species)) {
  print(ggplot(my.iris[my.iris$Species == this.species, ], aes(Sepal.Length, Sepal.Width)) + geom_point() + ggtitle(this.species))
}
dev.off()

for(counter in 1:100) {
  my.value <- counter * 2
  if(counter == 54) {
    stop("incomprehensible error message")
  }
}


```

4. Similar to what we did with the correlation power simulations, write code (including a loop) that creates a power simulation for comparing the means of two normally-distributed sample.
  - compare the sample with both `t.test()` and `wilcox.test()`
  - put the p-values from both tests in each simulation result
  - compare the power of the t-test vs. the power of the Wilcoxon test
  
```{r}
# set up the parameters of the simulation
n.sim <- 1000

# set up the "empty" data frame for the results
sim.results <- data.frame(sim = 1:

# start the loop
# each iteration of the loop should:
# 1. generate data
# 2. perform the analysis
# 3. store the results (here, p-values) in the results data frame


# examine the results
# use ifelse to create a column that indicates which of the p-values reached significance


# power = the proportion of "significant" results


```

