---
title: "Day 3 Practice"
output: html_document
---

1. Make a "copy" of the `iris` data set in your workspace. Use `summary()` on it, identify a factor column.

```{r}
my.iris <- iris
summary(my.iris)
```

Species is a factor

2. Create different versions of the factor column, in the new data set:
   a. a numeric vector with `as.numeric()`
   b. a character vector
   
```{r}
my.iris$SpeciesNum <- as.numeric(my.iris$Species)
my.iris$SpeciesChar <- as.character(my.iris$Species)
head(my.iris)
```

3. Use `xtabs()` to see how the values in these new columns line up with the original factor column.

```{r}
xtabs(~ Species + SpeciesNum, my.iris)
xtabs(~ Species + SpeciesChar, my.iris)
```

4. Create another column that is a copy of the factor column. Now, recode this new column so that the factor levels have new names. 

```{r}
my.iris$Species2 <- my.iris$Species
levels(my.iris$Species2) 
levels(my.iris$Species2) <- c("pretty", "ugly", "prettiest")
head(my.iris)
xtabs(~ Species + Species2, my.iris)
```

5. Check the order of levels of Species with `levels()`. Create a new column from this factor, changing the order of the levels, using:
  a. `relevel()`
  b. `reorder()`
  c. `factor()`

```{r}
levels(my.iris$Species)
my.iris$Species3 <- relevel(my.iris$Species, "virginica")
xtabs(~ Species + Species3, my.iris)

my.iris$Species4 <- reorder(my.iris$Species, my.iris$Sepal.Width)
xtabs(~ Species + Species4, my.iris)

my.iris$Species5 <- factor(my.iris$Species, levels = c("virginica", "versicolor", "setosa"))
xtabs(~ Species + Species5, my.iris)

str(my.iris$Species)
str(my.iris)
```

6. Working with missing values
   - read in the data from "scottdata/CognitionPaperModifiedData.csv"
   - locate NAs
   - display the rows of the data frame where the WithinMinPairs variable has missing values
   - create a new column that's a copy of the WithinMinPairs column
   - replace all missing values in this column with zeros
   - try using `xtabs()` to check this change in coding (why doesn't this work?)
   - try another method for double-checking your coding

```{r}

```

