---
title: "Day 4 Practice"
output: html_document
---

1. What are the four primary/basic data types in R?
   - Practice creating a vector of each data type
   - Using c(), concatenate pairs of these vectors together
   - What happens?  What data types are coerced to what?

```{r}
numbers <- 5:8
class(numbers)
doubles <- c(1.1, 73, 300, -5.45)
class(doubles)
typeof(doubles)
characters <- c("Alice", "Bob", "Charlie", "Daisy")
class(characters)
factors <- as.factor(c("Red", "Blue", "Green", "Yellow"))
class(factors)
levels(factors)
factors <- factor(c("Red", "Blue", "Green", "Yellow"), levels = c("Red", "Blue", "Green", "Yellow"))
levels(factors)
logicals <- c(TRUE, FALSE, FALSE, TRUE)
class(logicals)

c(numbers, doubles)
class(c(numbers, doubles))

c(numbers, characters)
class(c(numbers, characters))

x <- c(numbers, doubles)
c(x, characters)

c(numbers, factors)
class(c(numbers, factors))

c(doubles, factors)
class(c(doubles, factors))

c(characters, factors)
class(c(characters, factors))
c(characters, as.character(factors))

c(numbers, logicals)
class(c(numbers, logicals))

if(TRUE) print("yep, it's true")
if(FALSE) print("yep, it's true")
if(0) print("yep, it's true")
if(1) print("yep, it's true")
if(33) print("yep, it's true")
if(-33) print("yep, it's true")
if("Bob") print("yep, it's true")

c(characters, logicals)
class(c(characters, logicals))

c(factors, logicals)
class(c(factors, logicals))

c(numbers, c("1", "2", "3", "4"))
c(numbers, as.numeric(c("1", "2", "3", "4")))
as.numeric("Bob")

```

2. Make a copy of the `iris` data frame
  - Using `sample()` pull the data apart into two random batches
  - Put the pieces back together with `rbind()`
  - Put the pieces back together with `bind_rows()` from the `dplyr` package, creating an ID that keeps track of which piece is which
  
  
```{r}


```

3. Make a copy of the `iris` data frame
  - using [ , ] notation, pull the data frame into two sets of different columns
  - Put the pieces back together with `cbind()`
  - Put the pieces back together with `bind_cols()` from the `dplyr` package

```{r}


```

4. Run the following code to create a new small data frame with some additional information about iris species

```{r}
iris.info <- data.frame(Plant = c("setosa", "virginica", "limniris"),
                        State = c("Alaska", "Tennessee", "Louisiana"),
                        Rating = c(1, 5, 15))
```

Now merge this information with the `iris` data to create a new data frame. Using the `merge()` function, experiment with all = TRUE and all = FALSE. Also experiment with the `dplyr` join functions:
  - `inner_join()`
  - `full_join()`
  - `left_join()`
  - `right_join()`
  - `semi_join()`
  - `anti_join()`
  
```{r}
nrow(iris)
iris.merge1 <- merge(iris, iris.info) # default all = FALSE
nrow(iris.merge1)
iris.merge2 <- merge(iris, iris.info, by.x = "Species", by.y = "Plant") # default all = FALSE
nrow(iris.merge2)
summary(iris.merge2)

iris.merge3 <- merge(iris, iris.info, by.x = "Species", by.y = "Plant", all = TRUE) 
nrow(iris.merge3)
summary(iris.merge3)

iris.merge4 <- left_join(iris, iris.info, by = c("Species" = "Plant"))
iris.merge4$Species <- as.factor(iris.merge4$Species)
nrow(iris.merge4)
summary(iris.merge4)

```

