---
title: "Day 5 Homework"
output: html_document
---

0. Reload your data, and re-use the code you already have for getting things formatted nicely (columns renamed, factors relabeled, etc.)

```{r}
library(tidyverse)
scottdata <- read.csv("scottdata/CognitionPaperFinalData.csv")
```

1. Aggregate (e.g., get mean/median/max/something) of some numerical variable, over one or more factors.  Use `group_by()` plus `summarize()` from the `dplyr` package. You may or may not need to reshape the data first!

```{r}
# normal "nested" syntax
summarize(group_by(scottdata, Merged), 
          meanMP = mean(BetweenMinPairs), 
          medianMP = median(BetweenMinPairs), 
          sdMP = sd(BetweenMinPairs), 
          N = length(BetweenMinPairs), 
          seMP = sd(BetweenMinPairs)/sqrt(length(BetweenMinPairs)))

# "piped" syntax
group_by(scottdata, Merged) %>% summarize(meanMP = mean(BetweenMinPairs), 
                                          medianMP = median(BetweenMinPairs), 
                                          sdMP = sd(BetweenMinPairs), 
                                          N = length(BetweenMinPairs), 
                                          seMP = sd(BetweenMinPairs)/sqrt(length(BetweenMinPairs)))

```

2. Get your data into "long" (aka "tidy") format, if it isn't already

```{r}
# Already there!
```







