---
title: "Day 5 Practice"
output: html_document
---

1. Make another copy of the iris data set
  - Using `data.frame()` create a (small) table that assigns a color to each species.  That is, make a column called Species and a column called Color, with one row per species
  - use `merge()` to merge the data into the copy of the iris data set
  - create another table with an additional species (and color)
  - merge this new table with iris. What happened?
  - create one more table matching species to color, with only two rows, missing setosa
  - merge this new table with the copy of iris. What happened?

```{r}
library(dplyr)
my.iris <- iris
species.color <- data.frame(Species = c("setosa", "virginica", "versicolor"),
                            Color = c("red", "blue", "yellow"))
species.color
iris.merge1 <- merge(my.iris, species.color) #, by = "Species")
head(iris.merge1)
nrow(iris.merge1)
xtabs(~ Species + Color, iris.merge1)

species.color2 <- data.frame(Species = c("setosa", "virginica", "versicolor", "scorpiris"),
                            Color = c("red", "blue", "yellow", "purple"))
species.color2
iris.merge2 <- merge(my.iris, species.color2, all = FALSE)
iris.merge3 <- merge(my.iris, species.color2, all = TRUE)
iris.merge4 <- left_join(my.iris, species.color2)
iris.merge4$Species <- as.factor(iris.merge4$Species)
nrow(my.iris)
nrow(iris.merge2)
nrow(iris.merge3)
nrow(iris.merge4)

species.color3 <- data.frame(Species = c("virginica", "versicolor"),
                            Color = c("blue", "yellow"))
species.color3
iris.merge5 <- merge(my.iris, species.color3, all = FALSE)
iris.merge6 <- merge(my.iris, species.color3, all = TRUE)
iris.merge7 <- left_join(my.iris, species.color3)
iris.merge7$Species <- as.factor(iris.merge7$Species)
nrow(my.iris)
nrow(iris.merge5)
nrow(iris.merge6)
nrow(iris.merge7)

```

2. Make (yet another) copy of the iris data set
  - Create a column that takes the log of the Sepal.Width
  - Create a column that represents the multiplication of Sepal.Width by Sepal.Length
  - Create a "median split" column on Sepal.Width. In other words, calculate the median of Sepal.Width, and then create a column that indicates whether the value in that row is less than the median or not
  
```{r}
my.iris <- iris
my.iris$log.Sepal.Width <- log(my.iris$Sepal.Width)
my.iris$Sepal.Area <- my.iris$Sepal.Width * my.iris$Sepal.Length

median.Sepal.Width <- median(my.iris$Sepal.Width)

# one method
my.iris$median.split <- NA # not necessary (but maybe used to be?)
my.iris$median.split[my.iris$Sepal.Width < median.Sepal.Width]  <- "low"
my.iris$median.split[my.iris$Sepal.Width >= median.Sepal.Width] <- "high"

# another method
my.iris$median.split2 <- ifelse(test = my.iris$Sepal.Width < median.Sepal.Width, 
                               yes = "low", no = "high")
my.iris$median.split3 <- as.factor(ifelse(test = my.iris$Sepal.Width < median.Sepal.Width, 
                               yes = "low", no = "high"))


# another method
my.iris$median.split4 <- as.factor(my.iris$Sepal.Width < median.Sepal.Width)
levels(my.iris$median.split4) <- c("low", "high") # should be c("high", "low")

xtabs(~ Sepal.Width + median.split, my.iris)
xtabs(~ median.split + median.split2, my.iris)
xtabs(~ median.split + median.split3, my.iris)
xtabs(~ median.split + median.split4, my.iris)

```

3. Start with (yes, you guessed it!) another copy of the iris data set
  - Use `gather()` to convert this wide data to long format
    - You should be "gathering up" all of the columns with measurements
  - "Tidy" up the new factor column by splitting it into two columns: one designating dimension (width vs. length), one designating flower part (sepal vs. petal)
    
```{r}
my.iris <- iris
my.iris$PlantID <- as.factor(paste("plant", 1:nrow(my.iris), sep = ""))
iris.long <- gather(my.iris, measurement, cm, Sepal.Length:Petal.Width)
head(iris.long)

iris.long <- separate(iris.long, measurement, sep = "\\.", c("plant.part", "dimension"))
head(iris.long)
```

4. Continue working with the "long" iris data from previous problem
  - using `dplyr` functions, get the mean for each "cell" in the species x part x dimension design
  - make the results "wide", with different columns for each species

```{r}
cell.means <- group_by(iris.long, plant.part, dimension, Species) %>% summarize(mean.cm = mean(cm))
cell.means <- summarize(group_by(iris.long, plant.part, dimension, Species), mean.cm = mean(cm))

cell.means
cell.means.wide <- spread(cell.means, Species, mean.cm)
cell.means.wide

```

