---
title: "Day 5 Lecture"
output: html_document
---

# Outline
  - Review Homework 4
  - Merging with `merge()`
  - PRACTICE: merging
  - Manipulating and transforming variables
  - PRACTICE: transformations
  - Reshaping data & "tidy" data
  - SPLIT-APPLY-COMBINE
  - Aggregation (e.g., cells means)
  - PRACTICE: reshaping & aggregation
  - Pipe syntax from `magrittr`
  - Subsetting in `dplyr`
  - Complex objects
  - PRACTICE

# Review

# PRACTICE #1

# Data transformations    
## Creating new columns

```{r}
mysleep <- sleep
mysleep
mysleep$Study <- "Study1"
mysleep$Study2 <- c("Study1", "Study2")
mysleep$noise <- rnorm(20)
mysleep$noise2 <- rnorm(10)
#mysleep$noise2 <- rnorm(8)
mysleep
```

## Simple column transformations

```{r}
mysleep <- sleep
mysleep$extra2 <- mysleep$extra * 2
mysleep$extra3 <- log(mysleep$extra + 10)
mysleep
mysleep$extra2 <- mysleep$extra2 - 4
mysleep
mysleep$extra2 <- NULL  # deletes the column
mysleep
mysleep$extra3 <- NA
mysleep
mysleep$three <- 3
mysleep
# mysleep$four <- c(1, 2, 3) # doesn't work
mysleep$five <- c(1, 2, 3, 4) # does work (recycling)
mysleep

rep(2, times = 10)
rep(2:4, 10)
rep(2:4, each = 10)
rep(2:4, times = 3, each = 10)
rep(1:2, 3, each = 5)
c(rep(1:2, 3, each = 5), rep(1:8, 2, each = 3), 1:10)

```

# Keeping "tidy" data

    - `tidyr`
    - https://cran.r-project.org/web/packages/tidyr/vignettes/tidy-data.html
    - http://vita.had.co.nz/papers/tidy-data.html
    - Principle: variables/data in columns, observations/cases in rows
    - wide vs. long data
      - usually long data is easier to work with
      - sometimes wide formats are useful/convenient
      - moving between the two is important!
    - `reshape2` also very powerful package, if `tidyr` can't handle something

## Reshaping with `tidyr` 
   
```{r}
library(tidyr)
mysleep <- sleep
mysleep
levels(mysleep$group) <- c("pre", "post")

# spread(yourdata, factor.to.spread.by, column.with.values)

# QUICK PRACTICE: compute gains

# gather(yourdata, new.factor.column.name, measure.column.name, columns.to.gather)

```

## PRACTICE #2 & #3

## Splitting up columns
  - another "tidy data" goal
  - if a column represents multiple things/datapoints, split it up!
  - use `separate()` from `tidyr`

```{r}
colnames(who)
who.long <- gather(who, stuff, counts, new_sp_m014:newrel_f65)

who.long$stuff <- gsub("newrel", "new_rel", who.long$stuff)

who.long <- separate(who.long, stuff, sep = "_", c("newold", "type", "gender.age"))
summary(who.long)
head(who.long)

```

# Data cleaning
## A guideline template for data cleaning
  1. Get the data to read in to R
  2. Check all columns to make sure data is getting read in correctly
  3. Check NAs, decide what to do with them
  4. Remove excessive rows or other "defective" rows
  5. Reshape as needed (tend towards "long" formats)
  6. Get data out of column *names*, and into columns (variables)
  7. Fix/clean up column names
  8. Re-code factors as needed
  
## Reading data
  - find a good `read.X()` function, where X = your type of data
    - `read.table()` is the basic one
    - `read.csv()` and `read.delim()` are convenience "wrappers" for `read.table()`, for CSV and tab-delimited files, respectively
    - `read.spss()` and `read.dta()` from the `foreign` package for SPSS and Stata files
  - for messier or more primitive kinds of files in weird formats, use `readLines()`, or `scan()` to pull in the data in a very raw format, which you will then need to format in R

## Check all columns

## Check NAs
  - use summary()

## Remove rows/columns as needed

## Reshape as needed

## Get data out of column names, into variables

## Fix/improve column names

## Re-code factors as needed


# Split-apply-combine with the `dplyr` package
==============================================

## Aggregation

```{r}
library(dplyr)
library(reshape2)

myff <- french_fries
head(myff)

# make long
myff.long <- gather(myff, flavor, rating, potato:painty)
head(myff.long)

# get "cell means"
ff.means <- summarize(group_by(myff.long, treatment, flavor), mean = mean(rating))
ff.means
ff.means <- summarize(group_by(myff.long, treatment, flavor), mean = mean(rating, na.rm = TRUE))
ff.means

# other summary computations
ff.summaries <- summarize(group_by(myff.long, treatment, flavor), mean = mean(rating, na.rm = TRUE), std.dev = sd(rating, na.rm = TRUE), N = length(rating))
ff.summaries

```

## Forward-pipe operator (from `magrittr` package)
["This is not a pipe"](https://en.wikipedia.org/wiki/The_Treachery_of_Images)

  - `%>%` operator
  - effect: pass the results of the thing on the left to the first argument of the thing on the right
  - very commonly used in `dplyr`

```{r}
x <- 1:10
mean(log(x + 2))

(x + 2) %>% log() %>% mean()

summarize(group_by(myff.long, treatment, flavor), mean.rating = mean(rating, na.rm = TRUE))

# same as above, using the %>% operator
group_by(myff.long, treatment, flavor) %>% summarize(mean.rating = mean(rating, na.rm = TRUE))

```

## Subsetting with `dplyr`
  - `filter()` is for subsetting rows
  - `select()` is for subsetting columns

```{r}
sleep
sleep[sleep$group == 1, ]
filter(sleep, group == 1) # same results

sleep[sleep$group == 1 & sleep$extra > 0.2, ]
filter(sleep, group == 1, extra > 0.2) # same results

sleep[ , c("ID", "extra")]
select(sleep, ID, extra)
select(sleep, 3, 1)
select(sleep, c(3, 1))
#select(sleep, c("ID", "extra")) # doesn't work
select(sleep, c(ID, extra))
select_(sleep, "ID", "extra") # "underscore" variants of dplyr functions:
                              #   - can use strings to identify objects
                              #   - useful for more programmatic applications

```


# Notes on complex objects

  - use `names()` and `str()` to look at the structure of a complex object
  - most complex objects have a list-like structure
  - S3 vs. S4 classes
    - `$` notation works for S3
    - Try `@` instead for S4 classes
    - For more on S4: http://adv-r.had.co.nz/OO-essentials.html

```{r}
myff <- french_fries
results <- lm(potato ~ treatment, data = myff)
results
summary(results)
# plot(results)
class(results)
names(results)
str(results)
plot(results$residuals)
results.summary <- summary(results)
names(results.summary)
results.summary$adj.r.squared
results.summary$coefficients
class(results.summary$coefficients)
condition.pvals <- results.summary$coefficients[c("treatment2", "treatment3"), "Pr(>|t|)"]
condition.pvals

# str(summary(results))
# ?lm
```

The adjusted $R^2$ for this model is `r results.summary$adj.r.squared`.



